A scalar Poincaré map for anti-phase bursting in coupled inhibitory neurons with synaptic depression.

### 2022-03-05 tex/sections/results.tex
Edit introdu nullcline description

\documentclass[../manuscript.tex]{subfiles}
\graphicspath{{../figures/}}

\begin{document}

\subsection{Anti-phase burst solutions}
Short-term synaptic depression of inhibition in a half-centre oscillator acts as a \emph{burst termination} mechanism~\citep{brown1911} and is known to produce $n-n$ anti-phase burst solutions of varying period.
Such $n-n$ solutions consist of cells firing bursts of $n$ spikes in alternation.
\Cref{fig:depression-traces}A shows the timecourse of a typical $4-4$ burst.
While one cell is firing a burst it provides an inhibitory conductance to the other cell,  preventing it from firing.
Therefore, at any given moment one cell is spiking while the other is inhibited.
Consistent with~\citet{bose2011} we will refer to the currently firing cell as ``free'' and we will call the inhibited cell ``quiet''.
Additionally, we will distinguish between two phases of a $n-n$ solution:
We will refer to the burst duration of a cell as the ``free phase'', which is the time between the first spike and the last spike in a burst. And we will call the remaining duration of a cycle, when a cell is not spiking, the ``quiet phase''.

% TODO: Add mention of gstar. Once critical conductance is crossed, cell is released.
With each action potential of the free cell, short-term depression leads to a step-wise decrease of $d$, and consequently of $s$ (\cref{fig:depression-traces}B).
If $d$ depresses faster at spike time than it can recover in the inter-spike-intervals ($ISI$s), the total synaptic conductance $\gbar s$ will eventually become sufficiently small to allow for the quiet cell to be released and start firing, thus inhibiting the previously free cell.
While a cell is quiet its depression variable can recover.
Once the quiet cell becomes free again its synaptic inhibition will be sufficient to terminate the burst of the previously free cell and commence a new cycle.
As previously demonstrated by~\citet{bose2011}, in a two-cell reciprocally inhibitory network with synaptic depression the coupling strength $\gbar$ determines the type of $n-n$ solution.
Increasing $\gbar$ produces higher $n-n$ burst solutions with more spikes per burst and a longer cycle period.
\Cref{fig:burst-sols} shows numerically stable $n-n$ solutions for varying values of $\gbar$.
For small values of $\gbar$ the network produces anti-phase spiking $1-1$ solutions.
As $\gbar$ is increased the network generates solutions of increasing $n$, that is $2-2$, $3-3$, and $4-4$.
When $\gbar$ is sufficiently large (bottom of \cref{fig:burst-sols}), one of the cells continuously spikes at its uncoupled period $T$ while the other cell remains fully suppressed.
Depending on the initial conditions either of the two cells can become the suppressed cell, which is why the suppressed solution is numerically bistable.

Branches of numerically stable $n-n$ solutions and their associated limit cycle period for varying values of $\gbar$ are depicted in \cref{fig:bif-diagram}A (see  \textbf{Supplementary Material S2} for algorithm description).
Not only do higher $n-n$ solutions branches require stronger coupling $\gbar$, but also within $n-n$ branches the period increases with $\gbar$.
In line with~\citet{bose2011} we find small overlaps between solution branches indicating numerical bistability, for example such as between the $2-2$ and $3-3$ solution branches.
Branches of higher $n-n$ burst solutions occur on increasingly smaller intervals of $\gbar$, for instance is the $\gbar$ interval of the $5-5$ branch shorter than that of the $4-4$ branch and so on.
The interval between the $5-5$ branch and the suppressed solution (region between dotted lines in \cref{fig:bif-diagram}A) not only contains even higher numerically stable $n-n$ solutions, such as $11-11$ bursts, but also other non-symmetric $n-m$ solutions as well irregular, non-periodic solutions. However, the analysis in the following sections will only be concerned with the numerically stable and symmetric $n-n$ solutions.

\subsection{Mathematical analysis of two-cell network}
~\label{sec:assumptions}
% Summary of assumptions
The goal of the following mathematical analysis is to reduce the complexity of the eight-dimensional system to some easily tractable quantity.
As we will see later this quantity is the value of the depression variable $d$ of either of the two cells.
We will construct the solution of $d$ in a piecewise manner from one spike to the next, first during the free phase, and then during the quiet phase.
This construction will require two assumptions about the membrane and synaptic dynamics.
The first assumption states that during a burst the free cell fires at its uncoupled period $T$, which simplifies the construction of the solution of $d$.
The second assumption states that once the inhibitory conductance acting on the quiet cell drops below a critical threshold, the cell is immediately released and fires.
The second assumption is necessary to predict the release time of the quiet cell, which allows us to model the recovery of $d$ during the quiet phase.
In other words, the second assumption requires that the release of the quiet cell from inhibition depends only on the timecourse of the inhibition, and not on the membrane dynamics of the quiet cell.
Both assumptions can be observed in coupled relaxation-oscillator types of neurons such as the Morris-Lecar model we use, and will be numerically verified below.
Both assumptions were first explored in~\cite{bose2011} to derive algebraic conditions that guarantee the periodicity of the depression variable for different $n-n$ solutions.
However here we will use these assumptions to construct a Poincaré map of $d$, which will provide a geometric intuition for the dynamics of the full two-cell network and its dependence on model parameters.

% Assumption 1: Free phase
Our first assumption about the model states that the free cell fires at its uncoupled period $T$, that is, during the free phase of a burst we have $ISI=T$.
Solution profiles in \cref{fig:burst-sols} suggest that the $ISI$s are indeed approximately constant.
We can further numerically confirm this observation by capturing the $ISI$s of the stable solutions from the bifurcation diagram in \cref{fig:bif-diagram}A.
In addition to $ISI$s, \Cref{fig:bif-diagram}B also shows inter-\textit{burst} intervals ($IBI$s), which correspond to the time interval between the last spike of the burst and the full cycle period, and which lie in the quiet phase of the burst.
$IBI$s lie on multiple branches, each branch associated with a stable $n-n$ solution, and are monotonically increasing with $\gbar$.
In contrast, $ISI$s are calculated from the spikes within the free phase and do not vary significantly with $\gbar$, but are approximately $ISI\approx T$, affirming our first assumption.
Assuming $ISI=T$ allows us to ignore the non-linear membrane dynamics during the free phase, and to construct the evolution of the synaptic variables iteratively from spike to spike.
Assuming $ISI=T$ seems reasonable given that inhibition acting on the quiet cell decays exponentially to zero on a much shorter timescale than the duration of the $ISI$, and therefore, once the quiet cell is released its trajectory quickly approaches the spiking limit cycle.

% Assumption 2: Release
Our second assumption states that the quiet cell is released and spikes as soon as inhibition from the free cell drops below a constant threshold.
We will now define such a ``release condition'' by exploiting the discrepancy in timescales between the fast membrane dynamics, and the slower synaptic dynamics.
Let us first consider the dynamics of a single Morris-Lecar neuron.
We fix the synaptic variable that acts on the cell by setting $s=1$, which also makes the applied synaptic conductance $\gbar s$ constant.
Recall from \cref{fig:nullclines}A that in case of a single uncoupled cell ($\gbar s=0$), the $v$- and $w$-nullclines intersect at some unstable fixed point $p_{f}=(v_{f},w_{f})$, while trajectories revolve around a stable spiking limit cycle.
Increasing $\gbar s$ moves the cubic $v_{\infty}$ with the ensuing unstable fixed point $p_{f}$ down the sigmoid $w_{\infty}$ in the $(v-w)$-plane (\cref{fig:nullclines-coupled}).
When $\gbar s$ is large enough, the fixed point $p_{f}$ becomes stable, attracting all previously periodic trajectories.
There exists a unique value $\gbar s=g^{\star}$ when $p_{f}$ changes stability and the stable limit cycle vanishes.
Thus, when a constant inhibitory conductance is applied, $\gbar s<g^{\star}$ acts as a necessary condition for a cell to spike.
In contrast, when $\gbar s>g^{\star}$ inhibition is strong enough to prevent a cell from spiking~\citep{bose2011}.

@@

Now let us analyse the nullclines of the quiet cell when the two cells are coupled via synaptic inhibition with depression.
Let $\gbar s$ here denote the total synaptic conductance which acts on the quiet cell and is produced by the free cell, and let $p_{f}$ be the fixed point associated with the quiet cell.
At the start of the burst of the free cell we have $\gbar s > g^{\star}$ and $p_{f}$ is stable.
When the free cell spikes, $\gbar s$ peaks (\cref{eq:s-reset}), and the $v$-nullcline with the ensuing stable $p_f$ move down the $w$-nullcline.
Then in between spikes $\gbar s$ decays exponentially (\cref{eq:dot-s}) causing the $v$-nullcline and $p_{f}$ to move up the $w$-nullcline while attracting trajectories of the quiet cell.
Once depression causes the synaptic conductance to become small enough to satisfy $\gbar s<g^{\star}$ and the quiet cell is released, fixed point $p_{f}$ becomes unstable allowing the quiet cell to fire.
If the trajectory of the quiet cell remains sufficiently close to the stable $p_{f}$ at the time when it changes stability, then $\gbar s < g^{\star}$ acts as a release condition that is not only necessary, but also sufficient for firing of the quiet cell.
In this case the release of the quiet cell occurs precisely when
\begin{equation}
  ~\label{eq:release}
  \gbar s=g^{\star}
\end{equation}
is satisfied.

Whether the $(v,w)$-trajectory of the quiet cell can remain close enough to $p_{f}$ to make \cref{eq:release} sufficient for firing depends largely on the coupling strength $\gbar$ and the timescale disparity between membrane dynamics and synaptic dynamics~\citep{bose2011}.
It is straightforward to test our assumption of a release condition by numerically integrating the full system of ODEs and calculating the time interval between the first spike of the quiet cell and the time when $\gbar s$ first crosses $g^{\star}$.
We will call this time interval the ``release delay''.
If our assumption holds, we would expect an approximately zero release delay.
\Cref{fig:release-delay} shows the numerically computed graph of the release delay for varying $\gbar$.
The graph shows three distinct branches, next to each branch we also plot the timecourse of a corresponding sample solution of the total synaptic conductance $\gbar s$ of both cells.
For the rightmost branch where $\gbar>0.592$ $\si{mS/cm^{2}}$ the release delay is approximately zero.
Here the first spike of the quiet cell can be accurately predicted by the release condition in \cref{eq:release}.
The leftmost and middle branches for $\gbar<0.592$ $\si{mS/cm^{2}}$  show a release delay greater than zero, the quiet cell does not immediately fire when the release condition is first satisfied, and \cref{eq:release} does not accurately predict the release of the quiet cell.
The leftmost and middle branches contain $1-1$ and $2-2$ solutions respectively.
In both cases the coupling $\gbar$ is not sufficiently strong to allow trajectories of the quiet cell to be close enough to $p_{f}$ to guarantee spiking once $\gbar$ crosses $g^{\star}$.
Note that in the middle branch $\gbar s $ crosses $g^{\star}$ twice, and only after the second crossing does the quiet cell fire.
The following map construction relies on the assumption that the release condition in \cref{eq:release} can accurately predict the release time of the quiet cell.
Given our model parameters this is only possible for sufficiently large $\gbar>0.592 $ $\si{mS/cm^{2}}$.
For completeness, however, we will also consider values $\gbar < 0.592$ $\si{mS/cm^{2}}$ in the following analysis, bearing in mind that in this parameter range our map will not be accurate.

In summary: For $\gbar>0.592$ $\si{mS/cm^{2}}$ the release condition is sufficient to predict when the quiet cell is released.
Due to the symmetry of $n-n$ solutions the release occurs at exactly half the period of the full cycle, that is at $P/2$.
The release time therefore uniquely determines the type of $n-n$ solution.
Furthermore, computation of the release time does not depend on the membrane nor the synaptic dynamics of the quiet cell.
Instead, the solution of the synaptic variable $s$ of the free cell is sufficient to predict when $\gbar s=g^{*}$ is satisfied.
Finally, $s$ is solely determined by the evolution of the depression variable $d$ of the free cell.
Constructing a solution of $d$ during the free phase of either cell will therefore uniquely determine the solution of the full eight-dimensional network.
However, finding the solution $d$ requires us to know the initial value $d(0)$ at the start of a cycle at $t=0$.
In the next section we will construct a scalar return map that tracks these initial values $d(0)$ from cycle to cycle of stable $n-n$ solutions.

\subsection{Construction of the scalar Poincaré map}
In this section we construct the scalar Poincaré map $\Pi_{n}:d^{\star}\mapsto d^{\star}$.
Here the discrete variable $d^{\star}$ tracks the values of the continuous depression variable $d$ at the beginning of each $n-n$ burst.
The map $\Pi_{n}$ therefore describes the evolution of $d$, of either of the two cells, from the beginning of one cycle to the beginning of the next cycle.
To simplify the map construction we will assume that a free cell fires exactly $n$ times before it becomes quiet.
Later we will relax this assumption.
We will construct $\Pi_{n}$ by evolving $d$ first during the free phase and then during the quiet phase of the $n-n$ limit cycle.
First, let us give explicit definitions of the free and quiet phases.
A schematic illustration of both phases is given in \cref{fig:free-quiet1}.

Suppose that at $t=0$ cell 1 becomes free with some initial $d(0)$.
Cell 1 then fires $n$ spikes at the uncoupled period $T=T_{act}+T_{inact}$.
Let $s(t)$ and $d(t)$ be the corresponding solutions of the synaptic and depression variables of cell 1.
After $n$ spikes the total conductance $\gbar s(t)$ acting on the quiet cell 2 has decayed sufficiently to satisfy the release condition~\eqref{eq:release}, that is at some time $t=(n-1)T + \delt$, where $\Delta t < T_{inact}$, we have $\gbar s(t)=g^{\star}$.
Cell 2 is then released and prevents cell 1 from further spiking.
Here $\delt$ is the time between the last spike of cell 1 and the first spike of cell 2 ~\citep{bose2011}.
Once released, cell 2 also fires $n$ spikes until cell 1 becomes free once again at the cycle period.
Let $P_n$ denote the full cycle period of a $n-n$ solution:
\begin{equation}
  ~\label{eq:P}
  P_n = 2(n-1)T + 2\delt.
\end{equation}
We can now define the free and quiet phases of cell 1 explicitly. The free phase is the time interval between the first and last spikes of the burst, that is for time $0<t<(n-1)T$.
During the free phase of cell 1, the quiet cell 2 is inhibited sufficiently strong to prevent it from firing, hence $\gbar s > g^{\star}$.
The quiet phase of cell 1 is the remaining duration of the cycle when the cell is not firing, that is for $(n-1)T < t < 2(n-1)T + 2\delt$.

Note that only the quiet phase depends on $\delt$ which will play a central role in the construction of $\Pi_{n}$.
From \cref{eq:P} $\delt$ can be be computed as
\begin{equation}
  ~\label{eq:delta-t-P}
  \delt = \frac{1}{2}P_n - (n-1)T.
\end{equation}
\noindent
We can use \cref{eq:delta-t-P} and the numerically computed bifurcation diagram of the period for stable $n-n$ solutions in \cref{fig:bif-diagram}A to obtain the graph of $\delt$ as a function of $\gbar$ (\cref{fig:delta-t}).
Each continuous branch of $\delt$ is monotonically increasing and corresponds to a $n-n$ burst:
Stronger coupling $\gbar$ increases the total synaptic conductance $\gbar s$ that acts on the quiet cell, thus delaying its release.
It is easy to see that for any $n$-branch we have $\delt<T$:
Once $\delt$ crosses $T$, the free cell can ``squeeze in" an additional spike and the solutions bifurcate into a $(n+1)-(n+1)$ burst.

Distinguishing between the active and silent phases of a cycle allows us to describe the dynamics of the depression variable $d$ explicitly for each phase.
As can be seen from \cref{fig:free-quiet1}C, during the active phase $d$ depresses during the active phase of spikes and recovers during the inactive phases of spikes.
% TODO need to sort out terminology here, i.e. inactive vs quiet
In contrast, during the silent phase $d$ only recovers and does not depress.
Given the initial $d^{\star}=d(0)$ at the beginning of the cycle and the number of spikes in the free phase $n$, we can now construct the burst map $\Pi_{n}$.
The map
\begin{equation}
  \Pi_{n}(d^{\star})=Q_{n}\big(F_{n}(d^{\star}\big))
\end{equation}
\noindent
is a composition of two maps.
Map
\begin{equation}
  F_{n}:d^{\star}\mapsto \delt
\end{equation}
models the evolution of $d$ in the free phase.
$F_{n}$ takes an initial value $d^{\star}$ and calculates the inter-burst-interval $\delt$.
Map
\begin{equation}
  Q_{n}:\delt \mapsto d^{\star}
\end{equation}
models the recovery of $d$ in the quiet phase.
Given some $\delt$ map $Q_n$ computes $\dstar$ at the start of the next cycle.

Our aim in the following analysis is to elucidate the properties of $\Pi_{n}$ and to understand the structure of its parameter space by exploring how the stable and unstable fixed points of $\Pi_{n}$ are created.
To that effect it will be useful to include not only positive, but also negative values of $d^{\star}$ to the domain of $\Pi_{n}$.
But it is important to add that values $d^{\star}<0$ are biologically impossible as the depression variable models a finite pool of neurotransmitters, and therefore must be positive.
Because $\Pi_{n}$ maps first from $d^{\star}$ to $\delt$, and then back to $d^{\star}$, we will also consider negative values of $\Delta t$, interpreting them as $n-n$ solutions with partially overlapping bursts.
As will become evident, $\delt<0$ is only a formal violation of the biological realism of the map $\Pi_{n}$, as numerically stable $n-n$ solutions of the full system of ODEs only exist for $\Delta t>0$.

>>>>

\section{Results}
\subsection{Anti-phase burst solutions}
Short-term synaptic depression of inhibition in a half-centre oscillator acts as a \emph{burst termination} mechanism \citep{brown1911} and is known to produce \(n-n\) anti-phase burst solutions of varying period.
Such \(n-n\) solutions consist of cells firing bursts of \(n\) spikes in alternation.
\Cref{fig:depression-traces}A shows the timecourse of a typical \(4-4\) burst.
While one cell is firing a burst it provides an inhibitory conductance to the other cell,  preventing it from firing.
% FIXME: This isn't consistent with Bose, that needs to be mentioned
Therefore, at any given moment one cell is spiking while the other is inhibited. Consistent with \citet{bose2011} we will refer to the currently firing cell as ``\free{}'' and we will call the inhibited cell ``\quiet{}''.
Additionally, we will distinguish between two phases of a \(n-n\) solution:
We will refer to the time interval when a cell is firing as the ``\free{} phase'', and we will call the remaining duration of a cycle, when a cell is not firing, the ``\quiet{}  phase''.

% TODO: Add mention of gstar. Once critical conductance is crossed, cell is released.
With each action potential of the \free{} cell, short-term depression leads to a step-wise decrease of \(d\), and consequently of \(s\) (\cref{fig:depression-traces}B).
If \(d\) depresses faster at spike time than it can recover in the inter-spike-intervals (\(ISI\)s), the total synaptic conductance \(\bar g s\) will eventually become sufficiently small to allow for the \quiet{} cell to be released and start firing, thus inhibiting the previously \free{} cell.
While a cell is \quiet{} its depression variable can recover.
Once the \quiet{} cell becomes \free{} again its synaptic inhibition will be sufficient to terminate the burst of the previously \free{} cell and commence a new cycle.
% BUG: This is not 100% true, gbar AND the initial conditions uniquely determine the type of n-n pattern.
As previously demonstrated by~\citet{bose2011}, in a two-cell reciprocally inhibitory network with synaptic depression the coupling strength \(\bar g\) determines the type of \(n-n\) solution.
Increasing \(\bar g\) produces higher \(n-n\) burst solutions with more spikes per burst and a longer cycle period.
\Cref{fig:burst-sols} shows numerically stable \(n-n\) solutions for varying values of \(\bar g\).
For small values of \(\bar g\) the network produces anti-phase spiking \(1-1\) solutions.
As \(\bar g\) is increased the network generates solutions of increasing \(n\), that is \(2-2\), \(3-3\), and \(4-4\).
When \(\bar g\) is sufficiently large (bottom of \cref{fig:burst-sols}), one of the cells continuously spikes at its uncoupled period \(T\) while the other cell remains fully suppressed.
Depending on the initial conditions either of the two cells can become the suppressed cell, which is why the suppressed solution is numerically bistable.

Branches of numerically stable \(n-n\) solutions and their associated limit cycle period for varying values of \(\bar g\) are depicted in \cref{fig:bif-diagram}A (see \textbf{Supplementary Material S2} for algorithm description).
Not only do higher \(n-n\) solutions branches require stronger coupling \(\bar g\), but also within \(n-n\) branches the period increases with \(\bar g\).
In line with~\citet{bose2011} we find small overlaps between solution branches indicating numerical bistability, for example such as between the \(2-2\) and \(3-3\) solution branches.
Branches of higher \(n-n\) burst solutions occur on increasingly smaller intervals of \(\bar g\), for instance is the \(\bar g\) interval of the \(5-5\) branch shorter than that of the \(4-4\) branch and so on.
The interval between the \(5-5\) branch and the suppressed solution (region between dotted lines in \cref{fig:bif-diagram}A) not only contains even higher numerically stable \(n-n\) solutions, such as \(11-11\) bursts, but also other non-symmetric \(n-m\) solutions as well irregular, non-periodic solutions.
However, the analysis in the following sections will only be concerned with the numerically stable and symmetric \(n-n\) solutions.

\subsection{Mathematical analysis of two-cell network}
\label{sec:assumptions}
% Summary of assumptions
The goal of the following mathematical analysis is to reduce the complexity of the
eight-dimensional system to some easily tractable quantity.
As we will see later this quantity is the value of the depression variable \(d\) of either of the two cells.
We will construct the solution of \(d\) in a piecewise manner from one spike to the next, first during the \free{} phase, and then during the \quiet{} phase.
This construction will require two assumptions about the membrane and synaptic dynamics.
The first assumption states that during a burst the \free{} cell fires at its uncoupled period \(T\), which simplifies the construction of the solution of \(d\).
The second assumption states that once the inhibitory conductance acting on the \quiet{} cell drops below a critical threshold, the cell is immediately released and fires.
The second assumption is necessary to predict the release time of the \quiet{} cell, which allows us to model the recovery of \(d\) during the \quiet{} phase.
In other words, the second assumption requires that the release of the \quiet{} cell from inhibition depends only on the timecourse of the inhibition, and not on the membrane dynamics of the \quiet{} cell.
Both assumptions can be observed in coupled relaxation-oscillator types of neurons such as the Morris-Lecar model we use, and will be numerically verified below.
Both assumptions were first explored in~\cite{bose2011} to derive algebraic conditions that guarantee the periodicity of the depression variable for different \(n-n\) solutions.
However here we will use these assumptions to construct a Poincaré map of \(d\), which will provide a geometric intuition for the dynamics of the full two-cell network and its dependence on model parameters.

% Assumption 1: Free phase
% FIXME: Probably good idea to remove IBI reference here (and in figure).
Our first assumption about the model states that the \free{} cell fires at its uncoupled period \(T\), that is, during the \free{} phase of a burst we have \(ISI=T\).
Solution profiles in \cref{fig:burst-sols} suggest that the \(ISI\)s are indeed approximately constant.
We can further numerically confirm this observation by capturing the \(ISI\)s of the stable solutions from the bifurcation diagram in \cref{fig:bif-diagram}A.
In addition to \(ISI\)s, \Cref{fig:bif-diagram}B also shows inter-\textit{burst} intervals (\(IBI\)s), which correspond to the time interval between the last spike of the burst and the full cycle period, and which lie in the \quiet{} phase of the burst.
\(IBI\)s lie on multiple branches, each branch associated with a stable \(n-n\) solution, and are monotonically increasing with \(\bar g\).
In contrast, \(ISI\)s are calculated from the spikes within the \free{} phase and do not vary significantly with \(\bar g\), but are approximately \(ISI\approx T\), affirming our first assumption.
Assuming \(ISI=T\) allows us to ignore the non-linear membrane dynamics during the \free{} phase, and to construct the evolution of the synaptic variables iteratively from spike to spike.
Assuming \(ISI=T\) seems reasonable given that inhibition acting on the \quiet{} cell decays exponentially to zero on a much shorter timescale than the duration of the \(ISI\), and therefore, once the \quiet{} cell is released its trajectory quickly approaches the spiking limit cycle.

% Assumption 2: Release
Our second assumption states that the \quiet{} cell is released and spikes as soon as the total inhibitory conductance from the \free{} cell drops below a constant threshold.
We will now define such a ``release condition'' by exploiting the discrepancy in timescales between the fast membrane dynamics, and the slower synaptic dynamics.
Let us first consider the dynamics of a single Morris-Lecar neuron.
We fix the synaptic variable that acts on the cell by setting \(s=1\), which also makes the applied synaptic conductance \(\bar g s\) constant.
Recall from \cref{fig:nullclines}A that in case of a single uncoupled cell (\(\bar g s=0\)), the \(v\)- and \(w\)-nullclines intersect at some unstable fixed point \(p_{f}=(v_{f},w_{f})\), while trajectories revolve around a stable spiking limit cycle.
Increasing \(\bar g s\) moves the cubic \(v_{\infty}\) with the ensuing unstable fixed point \(p_{f}\) down the sigmoid \(w_{\infty}\) in the \((v-w)\)-plane (\cref{fig:nullclines-coupled}).
When \(\bar g s\) is large enough, the fixed point \(p_{f}\) becomes stable, attracting all previously periodic trajectories.
There exists a unique value \(\bar g s=g^{\star}\) when \(p_{f}\) changes stability and the stable limit cycle vanishes.
Thus, when a constant inhibitory conductance is applied, \(\bar g s<g^{\star}\) acts as a necessary condition for a cell to spike.
In contrast, when \(\bar g s>g^{\star}\) inhibition is strong enough to prevent a cell from spiking~\citep{bose2011}.

@@

Now let us analyse the nullclines of the \quiet{} cell when the two cells are coupled via synaptic inhibition with depression.
Let \(\bar g s\) here denote the total synaptic conductance which acts on the \quiet{} cell and is produced by the \free{} cell, and let \(p_{f}\) be the fixed point associated with the \quiet{} cell.
At the start of the burst of the \free{} cell we have \(\bar g s > g^{\star}\) and \(p_{f}\) is stable.
When the \free{} cell spikes, \(\bar g s\) peaks (\cref{eq:s-reset}), and the \(v\)-nullcline with the ensuing stable \(p_f\) move down the \(w\)-nullcline.
Then in between spikes \(\bar g s\) decays exponentially (\cref{eq:dot-s}) causing the \(v\)-nullcline and \(p_{f}\) to move up the \(w\)-nullcline while attracting trajectories of the \quiet{} cell.
Once depression causes the synaptic conductance to become small enough to satisfy \(\bar g s<g^{\star}\) and the \quiet{} cell is released, fixed point \(p_{f}\) becomes unstable allowing the \quiet{} cell to fire.
If the trajectory of the \quiet{} cell remains sufficiently close to the stable \(p_{f}\) at the time when it changes stability, then \(\bar g s < g^{\star}\) acts as a release condition that is not only necessary, but also sufficient for firing of the \quiet{} cell.
In this case the release of the \quiet{} cell occurs precisely when
\begin{equation}
	~\label{eq:release}
	\bar g s=g^{\star}
\end{equation}
is satisfied.

Whether the \((v,w)\)-trajectory of the \quiet{} cell can remain close enough to \(p_{f}\) to make \cref{eq:release} sufficient for firing depends largely on the coupling strength \(\bar g\) and the timescale disparity between membrane dynamics and synaptic dynamics~\citep{bose2011}.
It is straightforward to test our assumption of a release condition by numerically integrating the full system of ODEs and calculating the time interval between the first spike of the \quiet{} cell and the time when \(\bar g s\) first crosses \(g^{\star}\).
We will call this time interval the ``release delay''.
If our assumption holds, we would expect an approximately zero release delay.
\Cref{fig:release-delay} shows the numerically computed graph of the release delay for varying \(\bar g\).
The graph shows three distinct branches, next to each branch we also plot the timecourse of a corresponding sample solution of the total synaptic conductance \(\bar g s\) of both cells.
For the rightmost branch where \(\bar g>0.592\) \(\si{mS/cm^{2}}\) the release delay is approximately zero.
Here the first spike of the \quiet{} cell can be accurately predicted by the release condition in \cref{eq:release}.
The leftmost and middle branches for \(\bar g<0.592\) \(\si{mS/cm^{2}}\)  show a release delay greater than zero, the \quiet{} cell does not immediately fire when the release condition is first satisfied, and \cref{eq:release} does not accurately predict the release of the \quiet{} cell.
The leftmost and middle branches contain \(1-1\) and \(2-2\) solutions respectively.
In both cases the coupling \(\bar g\) is not sufficiently strong to allow trajectories of the \quiet{} cell to be close enough to \(p_{f}\) to guarantee spiking once \(\bar g\) crosses \(g^{\star}\).
Note that in the middle branch \(\bar g s \) crosses \(g^{\star}\) twice, and only after the second crossing does the \quiet{} cell fire.
The following map construction relies on the assumption that the release condition in \cref{eq:release} can accurately predict the release time of the \quiet{} cell.
Given our model parameters this is only possible for sufficiently large \(\bar g>0.592 \) \(\si{mS/cm^{2}}\).
For completeness, however, we will also consider values \(\bar g < 0.592\) \(\si{mS/cm^{2}}\) in the following analysis, bearing in mind that in this parameter range our map will not be accurate.

In summary: For \(\bar g>0.592\) \(\si{mS/cm^{2}}\) the release condition is sufficient to predict when the \quiet{} cell is released.
Due to the symmetry of \(n-n\) solutions the release occurs at exactly half the period of the full cycle, that is at \(P/2\).
The release time therefore uniquely determines the type of \(n-n\) solution.
Furthermore, computation of the release time does not depend on the membrane nor the synaptic dynamics of the \quiet{} cell.
Instead, the solution of the synaptic variable \(s\) of the \free{} cell is sufficient to predict when \(\bar g s=g^{*}\) is satisfied.
Finally, \(s\) is solely determined by the evolution of the depression variable \(d\) of the \free{} cell.
Constructing a solution of \(d\) during the \free{} phase of either cell will therefore uniquely determine the solution of the full eight-dimensional network.
However, finding the solution \(d\) requires us to know the initial value \(d(0)\) at the start of a cycle at \(t=0\).
In the next section we will construct a scalar return map that tracks these initial values \(d(0)\) from cycle to cycle of stable \(n-n\) solutions.

\subsection{Construction of the scalar Poincaré map}
In this section we construct the scalar Poincaré map \(\Pi_n:d^{\star}\mapsto d^{\star}\).
Here the discrete variable \(d^{\star}\) tracks the values of the continuous depression variable \(d\) at the beginning of each \(n-n\) burst.
The map \(\Pi_n\) therefore describes the evolution of \(d\), of either of the two cells, from the beginning of one cycle to the beginning of the next cycle.
To simplify the map construction we will assume that a \free{} cell fires exactly \(n\) times before it becomes \quiet{}.
We will construct \(\Pi_n\) by evolving \(d\) first during the \free{} phase and then during the \quiet{} phase of the \(n-n\) limit cycle.
First, let us give explicit definitions of the \free{} and \quiet{} phases.
A schematic illustration of both phases is given in \cref{fig:free-quiet1}.

Suppose that at \(t=0\) cell 1 starts \free{} with some initial \(d(0)\). Cell 1 then fires \(n\) spikes at the uncoupled period \(T=T_{act}+T_{inact}\).
Let \(s(t)\) and \(d(t)\) be the corresponding solutions of the synaptic and depression variables of cell 1.
After \(n\) spikes the total conductance \(\bar g s(t)\) acting on the \quiet{} cell 2 has decayed sufficiently to satisfy the release condition~\eqref{eq:release}.
That is at some time \(t=(n-1)T + T_{act} + \Delta t\), where \(\Delta t < T_{inact}\), we have \(\bar g s(t)=g^{\star}\).
Cell 2 is then released and prevents cell 1 from further spiking.
Here \(\Delta t\) is the time between the last spike of cell 1 and the first spike of cell 2 ~\citep{bose2011}.
Once released, cell 2 also fires \(n\) spikes until cell 1 becomes \free{} once again.
Let \(P_n\) denote the full cycle period of a \(n-n\) solution:
\begin{equation}
	~\label{eq:P}
	P_n = 2\Big( (n-1)T + T_{act} + \Delta t \Big).
\end{equation}
We can now define the \free{} and \quiet{} phases of cell 1 explicitly.
The \free{} phase is the interval that lasts from the first spike time up until the beginning of the inactive phase of the last spike, last spikes of the burst, that is for time \(0<t<(n-1)T + T_{act}\).
During the \free{} phase of cell 1, the \quiet{} cell 2 is inhibited sufficiently strong to prevent it from firing, hence \(\bar g s > g^{\star}\).
The \quiet{} phase of cell 1 is the remaining duration of the cycle when the cell is not firing, that is for \((n-1)T + T_{act} < t < P_n\).
Hence the \quiet{} phase lasts for \((n-1)T + T_{act} + 2\Delta t\).

Note that only the \quiet{} phase depends on \(\Delta t\), which will play a central role in the construction of \(\Pi_n\).
From \cref{eq:P} \(\Delta t\) can be computed as
\begin{equation}
	\label{eq:delta-t-P}
	\Delta t = \frac{1}{2}P_n - (n-1)T - T_{act}.
\end{equation}
We can use \cref{eq:delta-t-P} and the numerically computed bifurcation diagram of the period for stable \(n-n\) solutions in \cref{fig:bif-diagram}A to obtain the graph of \(\Delta t\) as a function of \(\bar g\) (\cref{fig:delta-t}).
Each continuous branch of \(\Delta t\) is monotonically increasing and corresponds to a \(n-n\) burst:
Stronger coupling \(\bar g\) increases the total synaptic conductance \(\bar g s\) that acts on the \quiet{} cell, thus delaying its release.
It is easy to see that for any \(n\)-branch we have \(\Delta t < T_{inact}\):
Once \(\Delta t\) crosses \(T_{inact}\), the \free{} cell can ``squeeze in" an additional spike and the solutions bifurcate into a \((n+1)-(n+1)\) burst.

Distinguishing between the \free{} and \quiet{} phases of a cycle allows us to describe the dynamics of the depression variable \(d\) explicitly for each phase.
As can be seen from \cref{fig:free-quiet1}C, during the \free{} phase \(d\) depresses when $v > v_\theta$ and recovers when $v < v_\theta$.
In contrast, during the \quiet{} phase \(d\) only recovers and does not depress.
Given the initial \(d^{\star} = d(0)\) at the beginning of the cycle and the number of spikes in the \free{} phase \(n\), we can now construct the burst map \(\Pi_n\).
The map
\begin{equation}
	\Pi_n(d^\star)=Q_n\big(F_n(d^\star\big))
\end{equation}
\noindent
is a composition of two maps. Map
\begin{equation}
	F_n:d^\star\mapsto \Delta t
\end{equation}
models the evolution of \(d\) in the \free{} phase.
\(F_n\) takes an initial value \(d^{\star}\) and calculates \(\Delta t\).
Map
\begin{equation}
	Q_n:\Delta t \mapsto d^{\star}
\end{equation}
models the recovery of \(d\) in the \quiet{} phase.
Given some \(\Delta t\) map \(Q_n\) computes \(d^\star\) at the start of the next cycle.

Our aim in the following analysis is to elucidate the properties of \(\Pi_n\) and to understand the structure of its parameter space by exploring how the stable and unstable fixed points of \(\Pi_{n}\) are created.
To that effect it will be useful to include not only positive, but also negative values of \(d^{\star}\) to the domain of \(\Pi_{n}\).
But it is important to add that values \(d^{\star}<0\) are biologically impossible as the depression variable models a finite pool of neurotransmitters, and therefore must be positive.
Because \(\Pi_{n}\) maps first from \(d^{\star}\) to \(\Delta t\), and then back to \(d^{\star}\), we will also consider negative values of \(\Delta t\), interpreting them as \(n-n\) solutions with partially overlapping bursts.
As will become evident, \(\delt<0\) is only a formal violation of the biological realism of the map \(\Pi_{n}\), as numerically stable \(n-n\) solutions of the full system of ODEs only exist for \(\Delta t>0\).
